---
title: "RepData_peerassignment_project"
author: "STS"
date: "29/08/2020"
output: html_document
---

Peer assignment Course project 1 
=================================
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)

```


## Loading and preprocessing the data

the data has the said number of rows (and columns) - 

```{r input file}
##Loading the data (i.e. read.csv())
##Process/transform the data (if necessary) into a format suitable for your analysis

input.file = '/Users/samarthts/Documents/scripts_R/R_git/Repro_res/RepData_assignment_STS/RepData_PeerAssessment1/'
inp = read.csv(paste0(input.file,'activity.csv' ), header = T,as.is = T)

#check dimensions to confirm number of rows
dim(inp)
```

Questions
============

## Q1. What is mean total number of steps taken per day?


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.
#inp1 = inp[!is.na(inp$steps),]

total_dailysteps=aggregate(x = inp$steps,by = list(inp$date), FUN = sum)



```



```{r}
#Make a histogram of the total number of steps taken each day
hist(total_dailysteps$x,main = 'Histogram of Total number of daily steps', 
     xlab = 'Total steps taken per day')
```


#### The mean and median of the total number of steps taken per day


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.
#inp1 = inp[!is.na(inp$steps),]

#mean_daily_steps=aggregate(x = inp$steps,by = list(inp$date), FUN = mean)

TD_mean = mean(total_dailysteps$x, na.rm=T)

#two decimal points only
TD_mean = format(round(TD_mean, 2), nsmall = 2)

```

### The mean number of steps taken per day is `r TD_mean` 


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.


#median_daily_steps=aggregate(x = inp$steps,by = list(inp$date), FUN = mean)
TD_median=median(total_dailysteps$x, na.rm = T)

#two decimal points only
TD_median = format(round(TD_median, 2), nsmall = 2)


```

### The median number of steps taken per day is `r TD_median` 


## Q2. What is the average daily activity pattern?

```{r time series plot}
#Make a time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

#remove NAs
inp1 = inp[!is.na(inp$steps),]

mean_daily_steps_2=aggregate(x = inp1$steps,by = list(inp1$interval), FUN = mean)

p <- ggplot(mean_daily_steps_2, aes(x=Group.1, y=x)) +
  geom_line() + 
  xlab("Intervals") +ylab("Steps")
p

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}

mean_daily_steps_2$Group.1[which(mean_daily_steps_2$x == max(mean_daily_steps_2$x))]

# #p <- ggplot(data, aes(x=day, y=value)) +
#   geom_line() + 
#   xlab("")
# p
```

## Imputing missing values

The number of missing value rows are - 

```{r missing values}
#Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs). 

sum(is.na(inp$steps))

```



```{r}
#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. Create a new dataset that is equal to the original dataset but with the missing data filled in.

#STRATEGY - median step count across all days for that particular missing 5 minute interval 

#inp1 = inp[!is.na(inp$steps),]

median_daily_steps_2=aggregate(x = inp1$steps,by = list(inp1$interval), FUN = median)

inp_missing=inp[is.na(inp$steps),]

#change names to merge
names(median_daily_steps_2) = c('interval','steps')

#merge 
inp_miss_imputed=merge(inp_missing, median_daily_steps_2,by = 'interval' ,all.x = T)

new_inp_imputed=inp_miss_imputed[c(4,3,1)]

names(new_inp_imputed) = names(inp)

# rbind both dataframes

new_inp=rbind(new_inp_imputed, inp1)

#sanity check

#sum(is.na(new_inp))
#dim(new_inp)

```
```{r}

#Make a histogram of the total number of steps taken each day 

total_dailysteps_imputed=aggregate(x = new_inp$steps,by = list(new_inp$date), FUN = sum)

hist(total_dailysteps_imputed$x)

```

### Calculate and report the new mean and median total number of steps taken per day. 


```{r}
##add code for mean and median
```


### The new mean number of steps taken per day is `r mean(total_dailysteps_imputed$x)` 

### The new median number of steps taken per day is `r median(total_dailysteps_imputed$x)` 



### Do these values differ from the estimates from the first part of the assignment? 

### Yes and table here - 

```{r}
#table
```


## Q3. What is the impact of imputing missing data on the estimates of the total daily number of steps?

- Fewer daily steps and left-skewed histogram 

- More intervals with low frequency step counts

-?

## Q4. Are there differences in activity patterns between weekdays and weekends?

- Yes,there are activity differences as seen in the panel plots below:

```{r}

#For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

#new_inp

#Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

new_inp$week_factor = weekdays(as.Date(new_inp$date))

new_inp$weekday_or_end = 'weekday'

new_inp$weekday_or_end[new_inp$week_factor == 'Sunday'| new_inp$week_factor == 'Saturday'] = 'weekend'


#create factor first?
new_inp$weekday_or_end =  as.factor(new_inp$weekday_or_end)

```

### Panel plot for weekdays and weekends 

```{r}

#Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:

mean_daily_steps_3=aggregate(x =new_inp$steps,by = list(new_inp$interval,new_inp$weekday_or_end ), FUN = mean)


p1 <- ggplot(mean_daily_steps_3, aes(x=Group.1, y=x,)) +
  geom_line() + 
  xlab("Intervals") +ylab("Steps")

p1+ facet_wrap( ~Group.2, nrow = 2)

```

