---
title: "RepData_peerassignment_project"
author: "STS"
date: "29/08/2020"
output: html_document
---

#Peer assignment Course project 1 
=================================
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)

```


## Loading and preprocessing the data

the data has the said number of rows (nad columns) - 

```{r input file}
##Loading the data (i.e. read.csv())
##Process/transform the data (if necessary) into a format suitable for your analysis

input.file = '/Users/samarthts/Documents/Coursera/DS_reproducibility /'
inp = read.csv(paste0(input.file,'activity.csv' ), header = T,as.is = T)

#check dimensions to confirm number of rows
dim(inp)
```

## Questions
============

## Q1. What is mean total number of steps taken per day?


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.
#inp1 = inp[!is.na(inp$steps),]

total_dailysteps=aggregate(x = inp$steps,by = list(inp$date), FUN = sum)


```

### The total number of steps taken per day is `r total_dailysteps` 


```{r}
#Make a histogram of the total number of steps taken each day
hist(total_dailysteps$x)
```


#### The mean and median of the total number of steps taken per day


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.
#inp1 = inp[!is.na(inp$steps),]

mean_daily_steps=aggregate(x = inp$steps,by = list(inp$date), FUN = mean)


```

### The mean number of steps taken per day is `r mean_daily_steps` 


```{r}
#For this part of the assignment, you can ignore the missing values in the dataset.
#inp1 = inp[!is.na(inp$steps),]

median_daily_steps=aggregate(x = inp$steps,by = list(inp$date), FUN = mean)


```

### The median number of steps taken per day is `r median_daily_steps` 

